---
title: "IST/CST Results"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
refined = 'intermediates'
private = 'private'
figure = 'working_figure'
```

```{r}

#Figure 3 IST Composition Heatmaps

library("readr")
library("tidyverse")
library("RColorBrewer")
library("pheatmap")
library("viridis")
library("ggplot2")
library("igraph")
library("slam")
library("scales")

#Name of IST/cluster column in metadata
tphe.cc <- "TPHE.IST"
ics.cc <- "ICS.IST"

#Read in sample metadata
pth = file.path(refined, 'dmn')
tphe.md <- read_delim(file.path(pth, "tphe_md.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
ics.md <- read_delim(file.path(pth, "ics_md.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)

rns <- tphe.md[[1]]
tphe.md <- tphe.md[ , 2:ncol(tphe.md)]
rownames(tphe.md) <- rns

rns <- ics.md[[1]]
ics.md <- ics.md[ , 2:ncol(ics.md)]
rownames(ics.md) <- rns

#Read in sample composition tables
tphe.cd4 <- read_delim(file.path(pth, "tphe_cd4_comp.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
tphe.cd8 <- read_delim(file.path(pth, "tphe_cd8_comp.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
ics.cd4 <- read_delim(file.path(pth, "ics_cd4_comp.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
ics.cd8 <- read_delim(file.path(pth, "ics_cd8_comp.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)

rns <- tphe.cd4[[1]]
tphe.cd4 <- tphe.cd4[ , 2:ncol(tphe.cd4)]
rownames(tphe.cd4) <- rns

rns <- tphe.cd8[[1]]
tphe.cd8 <- tphe.cd8[ , 2:ncol(tphe.cd8)]
rownames(tphe.cd8) <- rns

rns <- ics.cd4[[1]]
ics.cd4 <- ics.cd4[ , 2:ncol(ics.cd4)]
rownames(ics.cd4) <- rns

rns <- ics.cd8[[1]]
ics.cd8 <- ics.cd8[ , 2:ncol(ics.cd8)]
rownames(ics.cd8) <- rns

#Match metadata and composition data
tphe.md <- data.frame(tphe.md)
ics.md <- data.frame(ics.md)
tphe.cd4 <- data.matrix(tphe.cd4)
tphe.cd8 <- data.matrix(tphe.cd8)
ics.cd4 <- data.matrix(ics.cd4)
ics.cd8 <- data.matrix(ics.cd8)
tphe.cd4 <- tphe.cd4[ , (colnames(tphe.cd4) %in% rownames(tphe.md))]
tphe.cd8 <- tphe.cd8[ , (colnames(tphe.cd8) %in% rownames(tphe.md))]
ics.cd4 <- ics.cd4[ , (colnames(ics.cd4) %in% rownames(ics.md))]
ics.cd8 <- ics.cd8[ , (colnames(ics.cd8) %in% rownames(ics.md))]

#Construct properly formatted annotations, color scheme, and composition matrices for use in pheatmap
tphe.cd4.anno <- tphe.md[colnames(tphe.cd4), tphe.cc, drop=FALSE]
tphe.cd8.anno <- tphe.md[colnames(tphe.cd8), tphe.cc, drop=FALSE]
ics.cd4.anno <- ics.md[colnames(ics.cd4), ics.cc, drop=FALSE]
ics.cd8.anno <- ics.md[colnames(ics.cd8), ics.cc, drop=FALSE]
tphe.cd4.anno[[tphe.cc]] <- factor(tphe.cd4.anno[[tphe.cc]])
tphe.cd8.anno[[tphe.cc]] <- factor(tphe.cd8.anno[[tphe.cc]])
ics.cd4.anno[[ics.cc]] <- factor(ics.cd4.anno[[ics.cc]])
ics.cd8.anno[[ics.cc]] <- factor(ics.cd8.anno[[ics.cc]])
tphe.cd4.anno <- tphe.cd4.anno[order(tphe.cd4.anno[[tphe.cc]]), , drop=FALSE]
tphe.cd8.anno <- tphe.cd8.anno[order(tphe.cd8.anno[[tphe.cc]]), , drop=FALSE]
ics.cd4.anno <- ics.cd4.anno[order(ics.cd4.anno[[ics.cc]]), , drop=FALSE]
ics.cd8.anno <- ics.cd8.anno[order(ics.cd8.anno[[ics.cc]]), , drop=FALSE]
colors <- colorRampPalette(rev(RColorBrewer::brewer.pal(n=7, name="RdYlBu")), bias=3)(100)
tphe.cd4.mat <- tphe.cd4[, rownames(tphe.cd4.anno), drop=FALSE]
tphe.cd8.mat <- tphe.cd8[, rownames(tphe.cd8.anno), drop=FALSE]
ics.cd4.mat <- ics.cd4[, rownames(ics.cd4.anno), drop=FALSE]
ics.cd8.mat <- ics.cd8[, rownames(ics.cd8.anno), drop=FALSE]
tphe.cd4.mat <- t(apply(tphe.cd4.mat, 1L, scales::rescale))
tphe.cd8.mat <- t(apply(tphe.cd8.mat, 1L, scales::rescale))
ics.cd4.mat <- t(apply(ics.cd4.mat, 1L, scales::rescale))
ics.cd8.mat <- t(apply(ics.cd8.mat, 1L, scales::rescale))

#Make the heatmaps
pheatmap(mat = tphe.cd4.mat, color = colors, annotation_col = tphe.cd4.anno, cluster_rows = FALSE, cluster_cols = FALSE, show_colnames = FALSE, gaps_col = cumsum(unname(table(tphe.cd4.anno[[tphe.cc]]))))
pheatmap(mat = ics.cd4.mat, color = colors, annotation_col = ics.cd4.anno, cluster_rows = FALSE, cluster_cols = FALSE, show_colnames = FALSE, gaps_col = cumsum(unname(table(ics.cd4.anno[[ics.cc]]))))
pheatmap(mat = tphe.cd8.mat, color = colors, annotation_col = tphe.cd8.anno, cluster_rows = FALSE, cluster_cols = FALSE, show_colnames = FALSE, gaps_col = cumsum(unname(table(tphe.cd8.anno[[tphe.cc]]))))
pheatmap(mat = ics.cd8.mat, color = colors, annotation_col = ics.cd8.anno, cluster_rows = FALSE, cluster_cols = FALSE, show_colnames = FALSE, gaps_col = cumsum(unname(table(ics.cd8.anno[[ics.cc]]))))


```


```{r}

#Figure 3 IST Occurence Over PMA

library(readr)
library(ggplot2)
library(tidyverse)
library(ggbeeswarm)




# To get CGA at timepoints
timeline = read_csv(file.path('data', 'subject_timeline.csv')) %>% mutate(SampleID = str_c(Subject, '_',  `Sequence Num`))
subject = read_csv(file.path('data', 'subject_covariates.csv')) %>% mutate(GAB = 37 - preterm_weeks, BirthCohort = ifelse(preterm_weeks > 0, 'Pre-term', 'Full-term'))
timeline = left_join(timeline, subject)
# recode by mean CGA
recode_ist = function(tab){
  s = tab %>% group_by(IST) %>% summarize(mcga = mean(CGA)) %>% arrange(mcga)
  s = s %>% mutate(oIST = fct_reorder(factor(str_c('oIST_', seq_len(nrow(.)))), mcga), oIST_num = as.numeric(oIST))
  left_join(tab, s)
}

# ICS by Subject and timepoint
pth = file.path(refined, 'dmn')
tphe <- read_delim(file.path(pth, "tphe_md.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600) %>% 
  mutate(IST = gsub('TPHE IST ([0-9])', 'TPHE_\\1', `TPHE IST`)) %>% left_join(timeline)%>% select(Subject, CGA = cga, IST, GAB, Visit = `Sequence Num`, BirthCohort) %>% recode_ist

ics <- read_delim(file.path(pth, "ics_md.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)%>% 
  mutate(IST = gsub('ICS IST ([0-9])', 'ICS_\\1', `ICS IST`)) %>% left_join(timeline) %>% select(Subject, CGA = cga, IST, GAB,Visit = `Sequence Num`, BirthCohort) %>% recode_ist






# ics <- read_tsv(file.path(pth, "ics_basic.txt")) %>% recode_ist()
# tphe <- read_tsv(file.path(pth, "tphe_basic.txt")) %>% recode_ist()


ggplot(ics, aes(y=CGA, x=oIST, color=GAB )) + scale_color_gradient2(midpoint=37, low="red", mid="blue", high="darkblue", space ="Lab" ) + geom_quasirandom() + coord_flip()

```

These were reordered by average PMA in the version of this figure in the paper.

```{r}
summary(lm(CGA ~ IST, data = ics))
```


```{r}
ggplot(tphe, aes(y=CGA, x=oIST, color=GAB )) + scale_color_gradient2(midpoint=37, low="red", mid="blue", high="darkblue", space ="Lab" ) + geom_quasirandom() + coord_flip()


```

```{r}
summary(lm(CGA ~ IST, data = tphe))
```

## Co-occurance of t cell measures

```{r}
ics_tphe = bind_rows(ICS = ics, TPHE = tphe, .id = 'assay')
assays_avail = ics_tphe %>% group_by(Subject, Visit, BirthCohort) %>% summarize(assays = str_c(assay, collapse = '_')) %>% mutate(assays = factor(assays, levels = c('ICS', 'TPHE', 'ICS_TPHE')))

counts_by_subj = assays_avail %>% group_by(Subject, assays, BirthCohort) %>% summarize(n = n()) %>% ungroup() %>% arrange(Subject, desc(n))
# Take modal scenario if a subject had different assays available at different time points (uncommon)
counts_by_subj = counts_by_subj[!duplicated(counts_by_subj$Subject),]
assays_by_term = with(counts_by_subj, table(n, BirthCohort, assays))
ftab = ftable(assays_by_term, row.vars = c('n', 'BirthCohort'))
ftab

write.ftable(ftab, file.path(refined, 'assay_consort_alternative.txt'))
```

Number of subjects with 1, 2 or 3 samples of the various assays, stratified by Term.

```{r t_ist_traj, dev = c('png', 'pdf'), fig.width = 2, fig.height = 4}
ics_tphe = ics_tphe %>% mutate(Subjectf = fct_reorder(factor(Subject), GAB))
traj_plot = ggplot(ics_tphe, aes(y = Subjectf, x = CGA, fill = oIST_num)) + 
  geom_point(pch = 22) + scale_fill_distiller('IST', palette = 'GnBu') + facet_wrap(~assay) + 
  theme_minimal() + scale_y_discrete(breaks = NULL) + ylab("Subjects") + xlab('PMA') + geom_text(aes(label = oIST_num), size = 1.5) + theme(legend.position = 'bottom')
trajs = ics_tphe %>% group_by(assay) %>% do(plot = {
 out = traj_plot %+% .
  print(out)
  out
})

```

```{r}

#Figure 4 CST Composition Heatmap

library("readr")
library("tidyverse")
library("RColorBrewer")
library("pheatmap")
library("viridis")
library("ggplot2")
library("igraph")
library("slam")
library("scales")

CLUSTER_COLUMN <- "Renamed_CST"

md.rec <- read_delim(file.path(pth, "rec_basic.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
rns <- md.rec[[1]]
md.rec <- md.rec[ , 2:ncol(md.rec)]
rownames(md.rec) <- rns

md.nas <- read_delim(file.path(pth, "nas_basic.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
rns <- md.nas[[1]]
md.nas <- md.nas[ , 2:ncol(md.nas)]
rownames(md.nas) <- rns

genera.rec <- read_delim(file.path(refined, "REC_top_taxa.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
rns <- genera.rec[[1]]
genera.rec <- genera.rec[ , 2:ncol(genera.rec)]
rownames(genera.rec) <- rns

genera.nas <- read_delim(file.path(refined, "NAS_top_taxa.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
rns <- genera.nas[[1]]
genera.nas <- genera.nas[ , 2:ncol(genera.nas)]
rownames(genera.nas) <- rns

md.rec <- data.frame(md.rec)
md.nas <- data.frame(md.nas)

genera.rec <- data.matrix(genera.rec)
genera.nas <- data.matrix(genera.nas)

genera.rec <- genera.rec[ , (colnames(genera.rec) %in% rownames(md.rec))]
genera.nas <- genera.nas[ , (colnames(genera.nas) %in% rownames(md.nas))]

anno.rec <- md.rec[colnames(genera.rec), CLUSTER_COLUMN, drop=FALSE]
anno.nas <- md.nas[colnames(genera.nas), CLUSTER_COLUMN, drop=FALSE]

cga.rec <- md.rec[colnames(genera.rec), c(CLUSTER_COLUMN, "CGA"), drop=FALSE]
cga.nas <- md.nas[colnames(genera.nas), c(CLUSTER_COLUMN, "CGA"), drop=FALSE]

anno.rec[[CLUSTER_COLUMN]] <- fct_reorder(cga.rec[[CLUSTER_COLUMN]], cga.rec$CGA, mean)
anno.nas[[CLUSTER_COLUMN]] <- fct_reorder(cga.nas[[CLUSTER_COLUMN]], cga.nas$CGA, mean)

anno.rec <- anno.rec[order(anno.rec[[CLUSTER_COLUMN]]), , drop=FALSE]
anno.nas <- anno.nas[order(anno.nas[[CLUSTER_COLUMN]]), , drop=FALSE]

top25.rec <- head(names(rev(sort(rowSums(genera.rec)))), 25)
top25.nas <- head(names(rev(sort(rowSums(genera.nas)))), 25)

colors <- colorRampPalette(rev(RColorBrewer::brewer.pal(n=7, name="RdYlBu")), bias=3)(100)

mat.rec <- genera.rec[top25.rec, rownames(anno.rec), drop=FALSE]
mat.nas <- genera.nas[top25.nas, rownames(anno.nas), drop=FALSE]

mat.rec <- t(apply(mat.rec, 1L, scales::rescale))
mat.nas <- t(apply(mat.nas, 1L, scales::rescale))

pheatmap(mat = mat.rec, color = colors, annotation_col = anno.rec, cluster_rows = TRUE, cluster_cols = FALSE, show_colnames = FALSE, gaps_col = cumsum(unname(table(anno.rec[[CLUSTER_COLUMN]]))))
pheatmap(mat = mat.nas, color = colors, annotation_col = anno.nas, cluster_rows = TRUE, cluster_cols = FALSE, show_colnames = FALSE, gaps_col = cumsum(unname(table(anno.nas[[CLUSTER_COLUMN]]))))

```

```{r}

#Figure 4 CST Occurence Over PMA

library(readr)
library(ggplot2)
library(tidyverse)
library(ggbeeswarm)


recode_cst = function(tab){
  s = tab %>% group_by(CST) %>% summarize(mcga = mean(CGA)) %>% arrange(mcga)
  s = s %>% mutate(oCST = fct_reorder(factor(str_c('oCST_', seq_len(nrow(.)))), mcga), oCST_num = as.numeric(oCST))
  left_join(tab, s)
}

rec <- read_tsv(file.path(pth, "rec_basic.txt")) %>% recode_cst

nas <- read_tsv(file.path(pth, "nas_basic.txt")) %>% recode_cst

ggplot(rec, aes(y=CGA, x=Renamed_CST, color=gaBirth)) + scale_color_gradient2(midpoint=37, low="red", mid="blue", high="darkblue", space ="Lab" ) + geom_quasirandom() + coord_flip()
ggplot(nas, aes(y=CGA, x=Renamed_CST, color=gaBirth)) + scale_color_gradient2(midpoint=37, low="red", mid="blue", high="darkblue", space ="Lab" ) + geom_quasirandom() + coord_flip()


```

# Trajectories

```{r cst_traj, dev = c('png', 'pdf'), fig.width = 4, fig.height = 6}
rec_nas = bind_rows(REC = rec, NAS = nas, .id = 'assay') %>% rename(GAB = gaBirth)
rec_nas = rec_nas %>% mutate(Subjectf = fct_reorder(factor(Subject), GAB))
traj_plot = ggplot(rec_nas, aes(y = Subjectf, x = CGA, fill = oCST_num)) + 
  geom_point(pch = 22) + scale_fill_distiller('CST', palette = 'GnBu') + facet_wrap(~assay) + 
  theme_minimal() + scale_y_discrete(breaks = NULL) + ylab("Subjects") + xlab('PMA') + geom_text(aes(label = oCST_num), size = 1.5) + 
  theme(legend.position = 'bottom')
trajs2 = rec_nas %>% group_by(assay) %>% do(plot = {
  out = traj_plot %+% .
  print(out)
  out
})
```

```{r combined_traj, dev = c('png', 'pdf'), fig.width=8, fig.height=6}
cowplot::plot_grid(plotlist = trajs$plot, ncol = 2)
cowplot::plot_grid(plotlist = trajs2$plot, ncol = 2)
```

#Supplementary Figure 3 PCoA Plots

Runs in qiime.

```sh



qiime diversity core-metrics-phylogenetic --i-phylogeny nas_rooted_tree.qza --i-table nas_table.qza --p-sampling-depth 1200 --m-metadata-file nas_basic.txt  --output-dir nas_cda --p-n-jobs 24

qiime diversity core-metrics-phylogenetic --i-phylogeny rec_rooted_tree.qza --i-table rec_table.qza --p-sampling-depth 2250 --m-metadata-file rec_basic.txt  --output-dir rec_cda --p-n-jobs 24

```

```{r}

#Supplementary Figure 3 Axis Density by Term

library(readr)
library(ggplot2)

rec <- read_tsv(file.path(pth, "rec_basic.txt"))

rec$BirthCohort <- ifelse(rec$gaBirth >= 37, "Full term", "Preterm")

pc1 <- read_tsv(file.path('intermediates', "rec_pc1.txt"))

df <- merge(rec, pc1, by = "SampleID")

ggplot(data = df, aes(x = PC1, group = BirthCohort, fill = BirthCohort)) + geom_density(adjust=1.5, position="fill")

nas <- read_tsv(file.path(pth, "nas_basic.txt"))

nas$BirthCohort <- ifelse(nas$gaBirth >= 37, "Full term", "Preterm")

pc1 <- read_tsv(file.path('intermediates', "nas_pc1.txt"))

df <- merge(nas, pc1, by = "SampleID")

ggplot(data = df, aes(x = PC1, group = BirthCohort, fill = BirthCohort)) + geom_density(adjust=1.5, position="fill")

```

#Supplementary Figure 5 gCST 3 Time to Occurrence Based on Tphe IST at Discharge

This works exactly the same as the survival analysis script in `02_network_modeling`, except you can specify that it run for only a subset of CSTs and immune variables, and for all of those that are significant a plot is generated.  Optionally, you can save the fitted models and filtered input data for the chosen CST-immune subset.

See 02_network_modeling "time to occurrence"/survival analysis script for additional comments and description.


```{r}


#Start time
start.time <- Sys.time()

library(survival)
library(icenReg)
library(survminer)
library(dplyr)
library(data.table)
library(readr)

type <- "alt_surv"

sites = c("REC")

#Included for fig making
sigCSTs <- c("REC_3")

#Included for fig making
sigVars <- c("TPHE_IST_Disch")

#Included for fig making
fits.alt_surv <- list()

#Included for fig making
table.alt_surv <- list()

for (site in sites) {
  
  mapping <- read_tsv(file.path(refined, sprintf("%s_Surv_Mapping.txt", site)))
  
  raw_table <- read_tsv(file.path(refined, sprintf("%s_Surv_Input.txt", site)))
  
  mapping_length <- ncol(mapping)
  
  table_length <- ncol(raw_table)
  
  for (cst_i in 3:table_length) {
    
    tmp.time <- Sys.time()
    
    cst <- colnames(raw_table[,cst_i])
    
    #Included for fig making 
    if (!(cst %in% sigCSTs)) { next ; }
    
    #Included for fig making
    fits.alt_surv[[cst]] <- list()
    
    #Included for fig making
    table.alt_surv[[cst]] <- list()
    
    cst_in <- data.frame(raw_table[, 1:2], raw_table[, cst_i])
    
    cst_in <- cst_in[complete.cases(cst_in[,]),]
    
    var_names <- list()
    voi_names <- list()
    var_coeffs <- list()
    var_exp_coeffs <- list()
    variable_pvals <- list()
    k <- 1
    
    for (var_i in 4:mapping_length) {
      
      var <- colnames(mapping[,var_i])
      
      #Included for fig making  
      if (!(var %in% sigVars)) { next ; }
      
      mapping_in <- data.frame(mapping[, 1:3], mapping[, var_i])
      
      fac = FALSE
      
      working_table <- merge(mapping_in, cst_in, by = "Subject")
      
      working_table <- working_table[complete.cases(working_table[,]), ]
      
      if (typeof(working_table[[var]]) == "character") {
        
        fac = TRUE
        
        tmp_table <- as.data.frame(table(working_table[[var]]))
        
        rare_cats <- list()
        n <- 1
        
        for (cat_i in 1:nrow(tmp_table)) {
          
          if (tmp_table[[2]][cat_i] < 10) {
            
            rare_cats[[n]] <- levels(tmp_table[[1]])[cat_i]
            n <- n + 1
            
          }
          
        }
        
        working_table <- working_table[!(working_table[[var]] %in% rare_cats), ]
        
      }
      
      if ((nrow(working_table) < 20) || ((fac) && (nrow(table(working_table[[var]])) < 2))) {
        
        next
        
      }
      
      working_table$cst_obs <- working_table[[cst]]
      
      if (fac) {
        
        working_table$voi <- factor(working_table[[var]])
        
      } else {
        
        working_table$voi <- c(scale(working_table[[var]], center = TRUE, scale = TRUE))
        
      }
      
      working_table$GAB <- ((working_table$gaBirth)/37) - 1
      
      cst_surv <- Surv(working_table$PrevSampleDOL, working_table$cst_obs, type = "interval2")
      
      #Included for fig making
      fits.alt_surv[[cst]][[var]] <- ic_par(cst_surv ~ MOD + GAB + voi, data = working_table, model = "aft", dist = "loglogistic")
      
      #Included for fig making
      table.alt_surv[[cst]][[var]] <- working_table
      
      #Included for fig making
      next
      
      #Included for fig making
      if(!(dir.exists(fig_path <- file.path(figure, "cst_occurence_figs", type))))
      dir.create(fig_path)
      if (fac) {
        
        fac_lvls <- levels(as.data.frame(table(working_table[[var]]))[[1]])
        df.new <- data.frame(voi = fac_lvls, GAB = c(-.2, -.2), MOD = c("Caesarean_Section", "Caesarean_Section"))
        rownames(df.new) <- fac_lvls
      } else {
        df.new <- data.frame(voi = c(1, -1), GAB = c(-.2, -.2), MOD = c("Caesarean_Section", "Caesarean_Section"))
        rownames(df.new) <- c('High', 'Low')
      }
       pdf(file.path(fig_path, sprintf("%s_%s_plot.pdf", cst, var)))
        plot(fits.alt_surv[[cst]][[var]], df.new, xlab = "Day of Life")
        dev.off()
        
      
      #Included for fig making
      next
      
      fit <- ic_par(cst_surv ~ MOD + GAB + voi, data = working_table, model = "aft", dist = "loglogistic")
      
      fit_summary <- summary(fit)
      
      for (p in 5:nrow(fit_summary$summaryParameters)) {
        
        var_names[[k]] <- var
        voi_names[[k]] <- rownames(fit_summary$summaryParameters)[p]
        var_coeffs[[k]] <- fit_summary$summaryParameters[p, 1]
        var_exp_coeffs[[k]] <- fit_summary$summaryParameters[p, 2]
        variable_pvals[[k]] <- fit_summary$summaryParameters[p, 5]
        k <- k + 1
        
      }
      
      if(!(dir.exists(place <- file.path(refined, type))))
      dir.create(place)
      
      cat("Current Variable: ", file = file.path(place, sprintf("%s_results/%s_fit_log.txt", type, cst)), append = TRUE)
      cat(var, file = file.path(place, sprintf("%s_results/%s_fit_log.txt", type, cst)), append = TRUE)
      cat('\n', file = file.path(place, sprintf("%s_results/%s_fit_log.txt", type, cst)), append = TRUE)
      capture.output(summary(fit), file = file.path(place, sprintf("%s_results/%s_fit_log.txt", type, cst)), append = TRUE)
      cat('\n###\n###\n###\n\n', file = file.path(place, sprintf("%s_results/%s_fit_log.txt", type, cst)), append = TRUE)
        
    }
    
    #Included for fig making
    next
    
    adj_variable_pvals <- p.adjust(variable_pvals, method = "fdr")
    results <- cbind(var_names, voi_names, var_coeffs, var_exp_coeffs, variable_pvals, adj_variable_pvals)
    
    write.csv(results, file.path(place, sprintf("%s_results/%s_model_pvals.csv", type, cst)))
    
    cat("\nTime taken to complete last CST: ")
    cat(difftime(Sys.time(), tmp.time, units = "mins"))
    
  }
  
}

#Included for fig making
#save(fits.alt_surv, file = "selected_fitted_alt_survival_models.rda")
#save(table.alt_surv, file = "selected_fitted_alt_surv_dfs.rda")

#Print total runtime
cat("\nTotal runtime:")
cat(difftime(Sys.time(), start.time, units = "hours"))
cat("\n\n\n")

```