---
title: "Network modeling"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
refined = 'intermediates'
private = 'private'
data = 'data'
```

These analyses were used to generate the network plots in figure 6, as well as 

# Analysis of CST Occurrence vs Immunological Parameters - quasi-Poisson Duration

Iterates through every pairwise combination of CST and immunological parameter (either IST or metacluster abundance at one of the three timepoints) and tests for a significant associations between the immune parameter in adjusted and unadjusted models.

```{r, echo = FALSE, include = FALSE}
library(readr)
library(dplyr)
library(broom)
library(survival)
library(icenReg)

refined = 'intermediates'


# Take a big ol' table
# names of response columns and predictor columns

# Iterate through pairs of columns in ist_tbl and microbe_tbl, adjusting for extra_tbl
# All keyed by Subject
# null_formula should include all confounders
# regression_fcn(formula, data) returns an object that permits an anova test and tidy method
# result_id?
pairwise_regression = function(joined_tbl, ist_fields, microbe_fields, null_formula, regression_fcn, result_id){
  ist_tbl = joined_tbl[ist_fields]
  response_tbl = joined_tbl[microbe_fields]
  extra_tbl = joined_tbl[setdiff(names(joined_tbl), c(ist_fields, microbe_fields))]
  
  table_length <- ncol(response_tbl) 
  full_formula = update(null_formula, . ~ . + voi)
  
  
  if(! dir.exists(pth <- file.path(refined,  sprintf("%s_results", result_id))))
    dir.create(pth)
  
  all_results = list()
  i = 1
  
  warn_trapped_fcn = purrr::quietly(regression_fcn)
  for( cst_i in seq_len(table_length)){
    response_nm = names(response_tbl)[cst_i]
    for( voi_i in names(ist_tbl)){
      this_data = tibble(response = response_tbl[[cst_i]], voi = ist_tbl[[voi_i]]) %>% 
        bind_cols(extra_tbl) %>% na.omit()
      if(is.numeric(this_data$voi)){
        this_data$voi = scale(this_data$voi)
      }
      try({
      fit.null = warn_trapped_fcn(null_formula, this_data)
      fit.full = warn_trapped_fcn(full_formula, this_data)
      model_test = anova(fit.null$result, fit.full$result, test = "Chisq")
      full_results = tidy(fit.full$result) %>% mutate(response = response_nm,
                                               voi = voi_i, produced_warning = length(fit.full$warnings) > 0)
      anova_p = model_test$`Pr(>Chi)`[2]
      anova_result = tibble(term = 'anova', p.value = anova_p, response = response_nm, voi = voi_i)
      all_results[[i]] = bind_rows(full_results, anova_result)
      })
      i = i + 1
    }
  }
 bind_rows(all_results )
}

```

```{r, results = 'asis'}
to_fit = expand.grid(site = c('REC', 'NAS'), #microbiome sampling site
                     model_type = c('logit', 'duration', 'time_to_event'), #type of association. logit = odds of CST ever, duration = number of days in CST, time_to_event = rate of first entry to CST
                     specification = c('baseline', # baseline = MOD, GAB,
                                       'full'),# full = MOD, GAB, ABX, MILK
                     stringsAsFactors = FALSE
                     ) 

logit =  function(form_, data_) glm(form_, family = 'binomial', data = data_)
duration = function(form_, data_) glm(form_, family = 'quasipoisson', data = data_)
time_to_event = function(form_, data_){
  data_ = dplyr::filter(data_, !is.na(response))
  data_$response = Surv(data_$PrevSampleDOL, data_$response, type = "interval2")
  res = ic_par(form_, data = data_, model = "aft", dist = "loglogistic")
  res
}

knitr::kable(to_fit)
```

Scenarios to fit.

```{r}
nabx_polish = read_csv('data/antibiotic_exposure.csv') %>% 
  tidyr::pivot_wider(Subject, names_from = 'discharge', values_from = 'Number of systemic antibiotic') %>% 
  rename(abx_hospital = 'FALSE', abx_discharge = 'TRUE') %>%
  mutate(abx_hospital = (abx_hospital - mean(abx_hospital))/sd(abx_hospital), abx_discharge = abx_discharge- mean(abx_discharge))


milk_months = read_csv(file.path('intermediates', 'milk_subject.csv')) %>% select(Subject, milk_months)

hospital_humilk = read_csv('data/milk_hospital.csv') %>% rename(milk_perinatal = `Any Human Milk Perinatal`)
milk = full_join(milk_months, hospital_humilk) %>% mutate(milk_months = milk_months - mean(milk_months, na.rm = TRUE))
```
                                                          
Set up confounders (Antibiotics and Milk).  Both are centered.  Hospital antibiotics (days) are Z-scored so represent SD increases.


```{r}
fit_scenario_result = list()
for(i in seq_len(nrow(to_fit))){
  this_fit = to_fit[i,]
  ist_extra_tbl <- read_tsv(file.path(refined, sprintf("%s_Surv_Mapping.txt", this_fit$site)))
  
  # Get response data and
  if(this_fit$model_type == 'logit'){ # Logistic
     microbe_tbl <- read_tsv(file.path(refined, sprintf("%s_Logit_Input.txt", this_fit$site))) %>%
       mutate_at(.vars = vars(-Subject, -Total), .funs = function(x) x=='Yes')
  } else if(this_fit$model_type == 'duration'){  # Duration
    microbe_tbl =  read_tsv(file.path(refined, sprintf("%s_Duration_Input.txt", this_fit$site)))
    
  } else if(this_fit$model_type == 'time_to_event'){ #Survival
    microbe_tbl = read_tsv(file.path(refined, sprintf("%s_Surv_Input.txt",  this_fit$site)))
  }
  
  joined_tbl = left_join(microbe_tbl, ist_extra_tbl, by = 'Subject')
  regr_fcn = get(this_fit$model_type)
  
  # baseline
  this_formula = formula(response ~  MOD + gaBirth)
  if(this_fit$specification == 'full'){
    this_formula = update(this_formula, . ~ . +  milk_perinatal + milk_months + abx_hospital + abx_discharge)
  }
  
  # Total used as offset
  if(this_fit$model_type %in% c('logit', 'duration')){
    this_formula = update(this_formula, . ~. + offset(log(Total)))
    
  } else{
    ist_extra_tbl = left_join(microbe_tbl[c('PrevSampleDOL', 'Subject')], ist_extra_tbl, by = 'Subject')
    microbe_tbl = select(microbe_tbl, -PrevSampleDOL)
  }
  
  ist_fields = ist_extra_tbl %>% select(contains('TPHE'), contains('CD4'), contains('CD8'), contains('ICS')) %>% names()
  microbe_fields = microbe_tbl  %>% select(contains('REC'), contains('NAS')) %>% names()
  joined_tbl2 = joined_tbl %>%
    left_join(milk) %>%
    left_join(nabx_polish)
  
  fit_scenario_result[[i]] = pairwise_regression(joined_tbl2, ist_fields = ist_fields, microbe_fields = microbe_fields, null_formula = this_formula, regression_fcn = regr_fcn, do.call(paste, c(list(sep = '_'), this_fit)))
 
}

```

Fit all the scenarios.

# Top interactions

```{r}
library(ggplot2)
all_fits = bind_rows(fit_scenario_result, .id = 'scenario') %>% left_join(to_fit %>% mutate(scenario = as.character(seq_along(site))))

top = all_fits %>% filter(term == 'anova') %>% group_by(site, model_type) %>% arrange(p.value) %>% do(head(., n = 1))

top_coefs = semi_join(all_fits, top[c('scenario', 'voi', 'response')]) %>% anti_join(tibble(term = c('anova', '(Intercept)')))

clamp = function (x, modulus = 5) 
{
    x[x < -modulus] = -modulus
    x[x > modulus] = modulus
    x
}

ggplot(top_coefs, aes(x = term, y = estimate, ymin = estimate - std.error, ymax = estimate + std.error, color = clamp(-log10(p.value)))) + geom_pointrange() + facet_wrap(~model_type  + voi + response, scales = 'free') + coord_flip()
```


# Network plot figures

```{r networks}
sign_max = function(x){
  sign(x[which.max(abs(x))])
}

enmatrix = function(x, rownames_from){
  y = x[,setdiff(names(x), rownames_from)]
  y = as.matrix(y)
  rownames(y) = x[[rownames_from]]
  y
}

per_type_fdr = filter(all_fits, term == 'anova')  %>% group_by(model_type, specification)%>% mutate(fdr = p.adjust(p.value, 'fdr'))

per_type_sign = all_fits %>%  filter(stringr::str_detect(term, stringr::fixed('voi'))) %>% group_by(model_type, specification, voi, response) %>% summarize(sign_max = sign_max(estimate))

per_type_fdr = left_join(per_type_fdr, per_type_sign) %>% mutate(signed_fdr = clamp(-log10(fdr), 5) * sign_max, zeroed_fdr = signed_fdr * (fdr< .1))



per_type_nest = per_type_fdr %>% select(voi, response, zeroed_fdr, model_type, specification) %>% tidyr::nest(data = c(zeroed_fdr, voi, response))

per_type_nest$adj_matrix = purrr::map(per_type_nest$data, ~ tidyr::pivot_wider(.x, voi, names_from = response, values_from = zeroed_fdr) %>% enmatrix('voi'))

#Bipartite Graph of CST vs Immunome Results

#Load packages (Probably don't even need half of them - I tried a lot of things before I settled on this relatively simple solution.)
library(ggplot2)
library(network)
library(igraph)
library(sna)
library(ggnetwork)
library(GGally)
library(ergm)
library(intergraph)
library(RColorBrewer)
library(easyGgplot2)

#Define constants

set.seed(11)

rec_csts = c("REC_4", "REC_1", "REC_2", "REC_9", "REC_10", "REC_5", "REC_8", "REC_6", "REC_3", "REC_13", "REC_7", "REC_11", "REC_12")
nas_csts = c("NAS_4", "NAS_1", "NAS_2", "NAS_9", "NAS_10", "NAS_5", "NAS_8", "NAS_6", "NAS_3", "NAS_13", "NAS_7", "NAS_11", "NAS_12")
#tests = c("logit", "surv", "binom", "alt_binom")
tests = c("logit", "alt_surv")

figs <- list()

#Loop through the sites and make a figure for each as the target site (i.e. predicted by the CSTs of the other two sites)
for (test in tests) {

  #Read in the site-specific adjacency matrix specifying associations between taxa and the CSTs of the other body sites
  adj.matrix <- read.table(file = paste(workDir, sprintf("renamed_adjusted_%s_input.csv", test), sep = "/"), header= T, sep= ",", row.names= 1)
  
  #Use that adjacency matrix to construct a network object that is bipartite
  net.obj <- network(adj.matrix, matrix.type = "bipartite", ignore.eval = FALSE, names.eval = "weights")
  
  #Color the CST vertices according to the body site they represent
  network::set.vertex.attribute(net.obj, "color", ifelse(net.obj %v% "vertex.names" %in% rec_csts, "red", ifelse(net.obj %v% "vertex.names" %in% nas_csts, "blue", ifelse(grepl("TPHE", net.obj %v% "vertex.names"), "orange", "green"))))
  
  #Shape the vertices according to the visit they represent
  network::set.vertex.attribute(net.obj, "shape", ifelse(grepl("Birth", net.obj %v% "vertex.names"), 19, ifelse(grepl("Disch", net.obj %v% "vertex.names"), 17, ifelse(grepl("OneYear", net.obj %v% "vertex.names"), 15, 18))))
  
  #Make non-significant associations invisible
  network::set.edge.attribute(net.obj, "alpha", ifelse(net.obj %e% "weight" == 0.0, 0, 1))
  
  #Make a color gradient that will be used to color the edges
  rbPal <- colorRampPalette(c('darkblue', 'cyan', 'white', 'yellow', 'red'))
  
  #Map edge weights (signed log10 FDR adjusted p-vals) to colors in the gradient
  ecols <- rbPal(10)[as.numeric(cut(network::get.edge.attribute(net.obj, "weights"), breaks = 10))]
  
  figs[[test]] <- ggnet2(net.obj, color = "color", shape = "shape", edge.color = ecols, label = TRUE, label.size = 3, fontface = "bold", edge.size = 1, layout.par = list(niter = 1000, cell.jitter = 0.25))

}

#ggplot2.multiplot(nas_graph, rec_graph, thr_graph, cols = 2)

#ggsave(paste(workDir, sprintf("%s_graph.pdf", site), sep = "/"), width = 15, height = 15, units = "in")

```


# Targeted analysis of Alloiococcus abundance, Tphe5, and acute illness



```{r}
library("readr")
library("tidyverse")
library(lme4)
library(geepack)

#Read in mapping file with metadata including Alloiococcus abundance, Tphe5 at birth or discharge, and acute illness
md.nas <- read_delim(file.path(refined, "NAS_Focused_Mapping.txt"), "\t", escape_double = FALSE, trim_ws = TRUE, guess_max = 3600)
rns <- md.nas[[1]]
md.nas <- md.nas[ , 2:ncol(md.nas)]
rownames(md.nas) <- rns
#Turn it into a dataframe
md.nas <- data.frame(md.nas)
#Extract only post-discharge samples
md.post <- md.nas[which(md.nas$PostInitialDischarge == "Yes"), ]

#Center and scale numerical variables and convert categorical variables to factors
md.post$Subject <- factor(md.post$Subject)
md.post$GAB <- c(scale(md.post$gaBirth, center = TRUE, scale = TRUE))
md.post$nDOL <- c(scale(md.post$DOL, center = TRUE, scale = TRUE))
md.post$MOD <- factor(md.post$MOD)
md.post$IllnessVisit <- factor(md.post$IllnessVisit)
md.post$TPHE_5 <- factor(md.post$TPHE_5) #This is a binary variable with an affirmative value if the subject exhibited TPHE 5 at either birth or discharge

md.post$allo_freq <- round(md.post$Reads*md.post$Alloiococcus) #Convert Alloiococcus relative abundance to counts

#Test Alloiococcus and TPHE_5 as predictors of acute illness, both by themselves and jointly, controlling for confounders.
summary(glmer(IllnessVisit ~ Alloiococcus + nDOL + GAB + MOD + (1|Subject), data = md.post, family = binomial))
summary(glmer(IllnessVisit ~ TPHE_5 + nDOL + GAB + MOD + (1|Subject), data = md.post, family = binomial))
summary(glmer(IllnessVisit ~ Alloiococcus + TPHE_5 + nDOL + GAB + MOD + (1|Subject), data = md.post, family = binomial))

#Test TPHE_5 and acute illness as predictors of Alloiococcus abundance, both by themselves and jointly, controlling for confounders.
summary(geeglm(allo_freq ~ DOL*gaBirth + MOD + TPHE_5, id = Subject, corstr = "exchangeable", family = poisson(link = 'log'), data = md.post, offset = log(Reads)))
summary(geeglm(allo_freq ~ DOL*gaBirth + MOD + IllnessVisit, id = Subject, corstr = "exchangeable", family = poisson(link = 'log'), data = md.post, offset = log(Reads)))
summary(geeglm(allo_freq ~ DOL*gaBirth + MOD + TPHE_5 + IllnessVisit, id = Subject, corstr = "exchangeable", family = poisson(link = 'log'), data = md.post, offset = log(Reads)))

```